---
title: "HW_1_Template"
author: '*** Lina Cao , Thursdays 1:15PM ***'
date: "SDGB 7844; Prof. Nagaraja; Fall 2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

 The vectors `state.name`, `state.area`, and `state.region` are pre-loaded in R and contain US state names,  area (in square miles), and  region respectively. 
 
2a) Identify the data type for `state.name`, `state.area`, and `state.region`.

```{r}
# "typeof" is a function to determine the type or storage mode of the required object

typeof(state.name)
typeof(state.area)
typeof(state.region)

# we can also use "mode" or "class" to get the results
#    as shown in "help", "integer" or "double" in "typeof()" will be placed as "numeric"

mode(state.name)
mode(state.area)
mode(state.region)
class(state.name)
class(state.area)
class(state.region)

# from the result and the commonsense, use "is.xxx" to get the exact type of the data
# if the answer is TRUE, it is the correct type

is.character(state.name)
is.numeric(state.area)
is.factor(state.region)
```
The results shows all true.
The type of state.name is character.
The type of state.area is numeric.
The type of state.region is factor.

2b) What is the longest state name (including spaces)? How long is it?

```{r}
#  COUNT HOW LONG IS THE LONGEST STATE NAME  #
#   "nchar()" is used to count the number of characters (or bytes or Width), whose element is illustrated as "state.name[]".  Since we know state.name belongs to characters, we have to use "nchar" instead of "length".
# "max()" is to print the biggest one in the lengths of the state names

max(nchar(state.name))

#  PRINT THE LONGEST STATE NAME  #
#  we have the longest name width of the states, check where the state is in the vector state.name
#  use "==" to judge every element in state.name, whether they reach the requirement

      ## nchar(state.name)==max(nchar(state.name))

#  a series of TRUE/FALSE will be returned
#  extract those element with TRUE in former criterion, use "state.name[]"

state.name[nchar(state.name)==max(nchar(state.name))]

```

The longest state name have 14 characters (include space). There are two states have 14 characters. They are "North Carolina" and "South Carolina".

2c) Compute the average area of the states which contain the word "New" at the start of the state name.  Use the function `substr()`.
 
```{r}
# firstly, extract the first three letters from each state name

      ## substr(state.name[],1,3)

# we are required to find those start with "New", use "==" to check those reach the criterion

      ## substr(state.name[],1,3)=="New"

# third, sum all the areas from the states satisfied the condition, where "state.area[]" is used to extract area

      ## sum(state.area[substr(state.name,1,3)=="New"])

# then calculate the numbers that are summed up

      ## length(state.area[substr(state.name,1,3)=="New"])

# finally divide sum by length

sum(state.area[substr(state.name,1,3)=="New"])/length(state.area[substr(state.name,1,3)=="New"])
```

Use "substr()" to extract the first three letters of each state names, and choose those answer is ¡°New¡±. After that, extract the area of the states, and add them up.
The result of the states that meet the requirement is 47095.5 .


2d) Use the function `table()` to determine how many states are in each region.  Include the table in your solutions.    

```{r}
# simply use table() and list the results of how many states in each region
      
table(state.region)

# but it is not consist with the required table format
# copypaste the needed information to the following chart.

```

   Region      | Number of States
---------------| ----------------
Northeast      |        9
South          |        16
North Central  |        12
West           |        13


## Question 2

Computing the geometric mean of a numerical vector `x`.  

2a)  Using a `for` loop, write code to compute the geometric mean of a numeric vector `x`.  Make sure your code  (i) removes any `NA` values and (ii) prints an error message if there are any non-positive values in `x`.

```{r}
x <- c(NA,4,67,3)
x <- x[!is.na(x)] # remove all the NA in vector x
n=length(x)     # count the number involveed in calculation which is used as radical exponent

prod.x <- 1     # give product of x[i] the initial value 1
for(i in 1:n){  # loop repeated n times
  
  prod.x <- prod.x*x[i]  # multiply each new element to the former result
    
}   # end loop 
  #   use "if statement"" to compute total product
if (prod.x<=0){
  print("Error!!There is non-positive value in x")  
  #   print an error message if non-positive values exist
}else{
  print(prod.x^(1/n))  
  #   print the final geometric mean when the elements are all positive
} # end if/else


##### amended version  ####
x <- c(5,9,67,3)
x <- x[!is.na(x)] # remove all the NA in vector x
n=length(x)     # count the number involveed in calculation which is used as radical exponent

prod.x <- 1     # give product of x[i] the initial value 1
if (all(x>0)){
  for(i in 1:n){  # loop repeated n times
    prod.x <- prod.x*x[i]  # multiply each new element to the former result
  } # end loop 
  print(prod.x^(1/n))  
  # print the final geometric mean when the elements are all positive and with no NA
}else{
  print("Error!!There is non-positive value in x")
} #end if/else
   

```

1b) Test your code on the following cases and show the output: (i) {4, 67, 3}, (ii) {`NA`, 4, 67, 3}, (iii) {0, `NA`}, 6}, (iv) {67, 3, infinity}, and (v) {-infinity, 67, 3}.  

```{r}
#### (i) {4,67,3} ####

x <- c(4,67,3)
x <- x[!is.na(x)] # remove all the NA in vector x
n=length(x)     # count the number included in calculation which is used as radical exponent

prod.x <- 1     # give product of x[i] the initial value 1
for(i in 1:n){  # loop repeated n times
  
  prod.x <- prod.x*x[i]  # multiply each new element to the former result
    
}   # end loop 
  ### use "if statement"" to compute total product
if (prod.x<=0){
  print("Error!!There is non-positive value in x")  
  ### print an error message if non-positive values exist
}else{
  print(prod.x^(1/n))  
  ### print the final geometric mean when the elements are all positive
} # end if/else


#### (ii) {NA,4,67,3} ####

x <- c(NA,4,67,3)
x <- x[!is.na(x)] # remove all the NA in vector x
n=length(x)     # count the number included in calculation which is used as radical exponent

prod.x <- 1     # give product of x[i] the initial value 1
for(i in 1:n){  # loop repeated n times
  
  prod.x <- prod.x*x[i]  # multiply each new element to the former result
    
}   # end loop 
  #   use "if statement"" to compute total product
if (prod.x<=0){
  print("Error!!There is non-positive value in x")  
  #   print an error message if non-positive values exist
}else{
  print(prod.x^(1/n))  
  #   print the final geometric mean when the elements are all positive
} # end if/else

#### (iii) {0,NA,6}  ####

x <- c(0,NA,6)
x <- x[!is.na(x)] # remove all the NA in vector x
n=length(x)     # count the number included in calculation which is used as radical exponent

prod.x <- 1     # give product of x[i] the initial value 1
for(i in 1:n){  # loop repeated n times
  
  prod.x <- prod.x*x[i]  # multiply each new element to the former result
    
}   # end loop 
  #   use "if statement"" to compute total product
if (prod.x<=0){
  print("Error!!There is non-positive value in x")  
  #   print an error message if non-positive values exist
}else{
  print(prod.x^(1/n))  
  #   print the final geometric mean when the elements are all positive
} # end if/else

#### (IV) {67,3,Inf} ####

x <- c(67,3,Inf)
x <- x[!is.na(x)] # remove all the NA in vector x
n=length(x)     # count the number included in calculation which is used as radical exponent

prod.x <- 1     # give product of x[i] the initial value 1
for(i in 1:n){  # loop repeated n times
  
  prod.x <- prod.x*x[i]  # multiply each new element to the former result
    
}   # end loop 
  ### use "if statement"" to compute total product
if (prod.x<=0){
  print("Error!!There is non-positive value in x")  
  ### print an error message if non-positive values exist
}else{
  print(prod.x^(1/n))  
  ### print the final geometric mean when the elements are all positive
} # end if/else

#### (v) {-Inf,67,3} ####

x <- c(-Inf,67,3)
x <- x[!is.na(x)] # remove all the NA in vector x
n=length(x)     # count the number included in calculation which is used as radical exponent

prod.x <- 1     # give product of x[i] the initial value 1
for(i in 1:n){  # loop repeated n times
  
  prod.x <- prod.x*x[i]  # multiply each new element to the former result
    
}   # end loop 
  #   use "if statement"" to compute total product
if (prod.x<=0){
  print("Error!!There is non-positive value in x")  
  #   print an error message if non-positive values exist
}else{
  print(prod.x^(1/n))  
  #   print the final geometric mean when the elements are all positive
} # end if/else

```


## Question 3

Perfect numbers are those where the sum of the proper divisors (i.e., divisors other than the number itself) add up to the number.  For example, 6 is a perfect number because its divisors, 1, 2, and 3, when summed, equal 6.

3a) The following code was written to find the first 2 perfect numbers: 6 and 28; however, there are some errors in the code and the programmer forgot to add comments for readability.  Debug and add comments to the following.  

```{r}
### the corrected code is 

num.perfect<-2   #   set the number of perfect number that is required to output
count<-1         #   set the initial number of perfect number (the first one to be found)
iter<-2          #   start from 2 to find perfect number
while(count<=num.perfect){
divisor<-1
  for(i in 2:(iter-1)){
     if (iter%%i==0) {
       divisor<-c(divisor,i)
     } # end IF statement
  }#endf FOR loop
  if(sum(divisor)==iter){
    print(paste(iter,"is","a","perfect","number",sep=" "))
    count<-count+1
   }#endif
iter<-iter+1
}#endwhileloop

```
EXPLANATION:
1.the 2nd line of the code should be "count <- 1", indicating the "1st number perfect number", rather than the "zero" one.
2.in the for loop, i should be given as the format of "i in a:b", there should be an "in" between "i" and "2:(iter-1)".
3.the format of "if statement" is wrong, which should appear a "{" after the condition part.
4.in equavalence judgment, double equal sign should be used instead of only one
5.when it is going to print the result, "paste()" can only separate characters with "sep=",but cannot separate a string. 

num.perfect<-2
count<-0                      ### 1.it should be set as "1", indicating the "1st number "
iter<-2
while(count<=num.perfect){
divisor<-1
  for(i2:(iter-1)){                  ### 2.i should be given as "i in a:b" in for loop"
  if(iter%%i==0)divisor<-c(divisor,i)    ### 3.format of "if statements" is wrong
  }#end for loop                                     
if(sum(divisor)=iter){                 ### 4."==" should be used in equavalence judgment
  print(paste(iter,"is a perfect number",sep="")    ### 5. paste() can only separate characters with "sep=" 
  count<-count+1
}#end if
iter<-iter+1
}#end while loop


3b) Use the function `date()` at the start and at the end of your amended code.  Then compute how long the program takes to run (you can do this subtraction by hand).   Find the run time when you set `num.perfect` to 1, 2, 3, and 4.  Make a table of your results.  What are the first four perfect numbers?

```{r}
date()
num.perfect<-4   #   set the number of perfect number that is required to output
count<-1         #   set the initial number of perfect number (the first one to be found)
iter<-2          #   start from 2 to find perfect number
while(count<=num.perfect){
divisor<-1
  for(i in 2:(iter-1)){
     if (iter%%i==0) {
       divisor<-c(divisor,i)
     } # end IF statement
  }#endf FOR loop
  if(sum(divisor)==iter){
    print(paste(iter,"is","the","perfect","number",sep=" "))
    count<-count+1
   }#endif
iter<-iter+1
}#endwhileloop)
date()
```

ENTER YOUR EXPLANATION FOR QUESTION 3b HERE.

Count | Start Time |  Stop Time | Elapsed Time (sec.)
----- | ---------- |  --------- | -------------------
  1   |  10:40:54  |  10:40:54  |     almost os
  2   |  10:39:36  |  10:39:37  |        1s
  3   |  10:41:37  |  10:41:38  |        1s
  4   |  10:42:00  |  10:42:10  |        10s
  
The first four perfect numbers are 6, 28, 496, 8128


3c) Let `x <- 1:4` and define `y` to be the vector of run times.  Plot `y` vs `x` using the code below.  Is the relationship between the discovery of perfect numbers and run times on your computer linear?  Justify your answer.

```{r}
x<-1:4
y<-c(0,1,1,10)                  
plot(x,y,pch=20,type="b",
     xlab="number of perfect numbers discovered",
     ylab="cumulative length of time(in seconds)",
     main="Cumulative Run Times to Discover Perfect Numbers",
     las=TRUE)
```

I do not think the relationship is linear. As is illustrated in the plot, the time to compute one, two and three perfect numbers are almost the same, while it costs much longer time when it comes to get four perfect numbers.
I tried to compute five perfect numbers, however, the fifth perfect number did not show up even after 15 minutes run. Therefore, I can simply conclude that the time will balloon exponentially as the number of required perfect number increase from four to five.




